---
title: "DATA358 Final"
author: "Taylor Stacy"
date: "`r Sys.Date()`"
output:
  html_document:
    number_section: no
    toc: no
    toc_depth: 3
    toc_float: yes
    code_folding: show
    css: lab_templet.css
---

```{r setup, include=FALSE}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(rsample)
library(MASS) 
library(caret)
library(reshape2)
library(here)
library(gridExtra)
library(factoextra)
library(kableExtra)
library(jtools)
library(readr)
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

## Data
```{r}
dat <- read_csv(here("Final Project/data358_finaldata_va.csv"))
summary(dat)
```
## Data Cleaning
### Missing Values
```{r}
dat %>% 
  summarise_all(funs(sum(is.na(.)))) %>% 
  gather()%>%
  mutate(percent=round(value/sum(value)*100,2))
datx <- dat %>%
  mutate(OffAge= ifelse(is.na(OffAge), mean(OffAge, na.rm=TRUE), OffAge)) %>%
  mutate(VicAge= ifelse(is.na(VicAge), mean(VicAge, na.rm=TRUE), VicAge)) %>%
  ungroup()
dat %>% 
  group_by(VicSex) %>% 
  summarize(count= n()) %>% 
  mutate(prop=round(count/sum(count),2)) %>% 
  arrange(desc(count))
dat %>% 
  group_by(VicRace) %>% 
  summarize(count= n()) %>% 
  mutate(prop=round(count/sum(count),2)) %>% 
  arrange(desc(count))
dat %>% 
  group_by(OffSex) %>% 
  summarize(count= n()) %>% 
  mutate(prop=round(count/sum(count),2)) %>% 
  arrange(desc(count))
dat %>% 
  group_by(OffRace) %>% 
  summarize(count= n()) %>% 
  mutate(prop=round(count/sum(count),2)) %>% 
  arrange(desc(count))
dat %>% 
  group_by(Weapon) %>% 
  summarize(count= n()) %>% 
  mutate(prop=round(count/sum(count),2)) %>% 
  arrange(desc(count))
datx <- datx %>% 
  mutate(VicSex= ifelse(is.na(VicSex),"Male", VicSex)) %>% 
  mutate(VicRace= ifelse(is.na(VicRace),"Black", VicRace)) %>% 
  mutate(OffSex= ifelse(is.na(OffSex), "Male" , OffSex)) %>% 
  mutate(OffRace= ifelse(is.na(OffRace),"Black", OffRace)) %>%
  mutate(Weapon = ifelse(is.na(Weapon), "Firearm", Weapon))
```

```{r}
finaldat <- datx %>% 
  mutate(Solved = as.factor(Solved)) %>% 
  mutate(Homicide = as.factor(Homicide)) %>% 
  mutate(Situation = as.factor(Situation)) %>% 
  mutate(VicSex = as.factor(VicSex)) %>% 
  mutate(VicRace= as.factor(VicRace)) %>% 
  mutate(OffSex = as.factor(OffSex)) %>% 
  mutate(OffRace= as.factor(OffRace)) %>% 
  mutate(Weapon = as.factor(Weapon))

finaldat %>% 
  summarise_all(funs(sum(is.na(.)))) %>% 
  gather()
```
```{r}
summary(finaldat)
```

## Distribution of Victims and Officers by Race and Sex
```{r}
ggplot(data = finaldat, aes(x=VicRace, fill = VicSex), horizontal = TRUE)+
  geom_histogram(stat = "count")+
  coord_flip()+
  theme_classic()
ggplot(data = finaldat, aes(x=OffRace, fill = OffSex))+
  geom_histogram(stat = "count")+
  coord_flip()+
  theme_classic()
```

The races white and black are the highest values for both the officers and victims, this can be explained by the US' population distribution, which has been predominately white and black. 

Most officers and victims' ethnicity was not reported or unknown. 

```{r}
ggplot(data = finaldat, aes(x=Weapon, fill = Weapon))+
  geom_histogram(stat = "count", binwidth=1,boundary=-0.5) +
  theme(text = element_text(size=10), axis.text.x = element_text(angle=45, hjust = 1))
```

## Data Splitting
```{r}
set.seed(1234)
inTrain <- createDataPartition(finaldat$Solved,
                               p=0.8,
                               list = FALSE)
train <- finaldat[inTrain,]
test <- finaldat[-inTrain,]
```
# Hierarchical Tree
## Distance Matrix
```{r}
start <- Sys.time()

res.dist <- dist(finaldat,
                 method = "euclidian")
fviz_dist(res.dist,
          lab_size = 2)

end <- Sys.time()
end - start
```

## Denrograph
```{r}
res.hc.ward <- hclust(
  d= res.dist,
  method = "ward.D2")
plot(res.hc.ward, hang = -1, cex = 0.6)
```

### Visualize Tree
```{r}
dat_num <- finaldat %>% 
  mutate(Solved = as.numeric(Solved)) %>% 
  mutate(Homicide = as.numeric(Homicide)) %>% 
  mutate(Situation = as.numeric(Situation)) %>% 
  mutate(VicSex = as.numeric(VicSex)) %>% 
  mutate(OffSex = as.numeric(OffSex)) %>% 
  mutate(VicRace = as.numeric(VicRace)) %>% 
  mutate(OffRace = as.numeric(OffRace)) %>% 
  mutate(Weapon = as.numeric(Weapon))

grp <- cutree(res.hc.ward, k = 4)
fviz_cluster(
  list(data = dat_num, cluster = grp),
  ellipse.type = "convex", 
  show.clust.cent = FALSE, 
  ggtheme = theme_minimal())
```
